---
title: "kaggle heart DS"
author: "shuo wang"
date: "06/11/2021"
output: pdf_document
---

```{r}
library(kableExtra)
library(dplyr)
library(tidyverse)
library(forcats)
heart_tbl<-read_csv("heart.csv")
glimpse(heart_tbl)
heart_tbl %>% arrange(Age) %>% head(10) %>% kable %>% kable_styling()
```

a\. Using the plot(s) of your choice, assess whether there is an association between the sex of the patient and their resting blood pressure, i.e. is there a difference in distribution of the resting heart rates across the sexes? Explain your answer.

ans: from the box-plot below, there is no significant association between the sex of the patient and their resting blood pressure

```{r}
 A<-heart_tbl %>% select(Sex, RestingBP) %>% drop_na()
ggplot(data = A, aes(x=Sex,y=RestingBP)) + geom_boxplot(aes(fill=Sex))

```

b\. Produce a stacked barplot showing the distribution of Chest Pain Type for each level of RestingECG.

```{r}
B<- heart_tbl %>% group_by(ChestPainType) %>% select(ChestPainType, RestingECG)
ggplot(data=B, mapping=aes(x=RestingECG, fill=ChestPainType))+geom_bar()+labs(title="Chest Pain distribution")

```

c\. Produce a summary table containing counts and proportions of RestingECG category for each sex/ChestPainType factor combination.

```{r}
heart_tbl %>% group_by(RestingECG) %>% count(Sex, ChestPainType, sort = TRUE) %>% mutate(proportion = n/sum(n)) %>% arrange(RestingECG) %>%  kable %>% kable_styling()
```

d.Create a summary table that finds the **mean, median and IQR** of RestingBP, Cholesterol, FastingBS, and MaxHR for each of the Chest Pain Types and report those results in a tibble where the columns are the levels of Chest Pain Types and the summary statistics are in the rows.

```{r}
D<- heart_tbl %>% group_by(ChestPainType) %>%select(RestingBP, Cholesterol,FastingBS,MaxHR) %>% summarise_all(
  list(mean = ~mean(.), median = ~median(.), IQRs = ~IQR(.)),
  na.rm=TRUE) 
 t(D) %>% kable %>% kable_styling()

```

e.Using the plot(s) of your choice, explain which of the following measurements seem most strongly associated with Heart Disease (heart disease vs. normal) : RestingBP, Cholesterol, FastingBS, and MaxHR.

**ANS:Fasting blood sugar seems have the most strong association with Heart Disease**

```{r}
ggplot(data = heart_tbl, aes(x=HeartDisease,y=RestingBP,group=HeartDisease)) + geom_boxplot(aes(fill=HeartDisease))
ggplot(data = heart_tbl, aes(x=HeartDisease,y=Cholesterol,group=HeartDisease)) + geom_boxplot(aes(fill=HeartDisease))
ggplot(data = heart_tbl, aes(x=HeartDisease,y=FastingBS,group=HeartDisease)) + geom_boxplot(aes(fill=HeartDisease))
ggplot(data = heart_tbl, aes(x=HeartDisease,y=MaxHR,group=HeartDisease)) + geom_boxplot(aes(fill=HeartDisease))

```

f.Create both a 2-d histogram and a 2-d contour plot to assess the association betwen RestingBP and MaxHR. Describe this association and also explain which plot you think shows the association most clearly (or explain why they are about the same).

**ANS: the contour plot shows more clearly that most people with MAX heart rate around 125 to150 ,RestingBP 125 to 140**

```{r}
ggplot(heart_tbl,aes(x=RestingBP, y=MaxHR))+geom_bin2d(bins = 50 )+theme_bw()
ggplot(heart_tbl,aes(x=RestingBP, y=MaxHR))+stat_density_2d(aes(fill = ..level..), geom = "polygon", colour="white")+ylim(c(0,200))+xlim(c(0,200))

```

g.Using the plot(s) of your choice, determine whether the association in (f) depends on either the Chest Pain Type or the Heart Disease status (or both).

**ANS: from plot2, people who have heart disease have higher max heart rate. however, in plot1, cant find significant influence of the association with different type of chest pain.**

```{r}
ggplot(data = heart_tbl, aes(x=RestingBP,y=MaxHR,group=ChestPainType,col=ChestPainType)) + geom_point()+ facet_wrap(~ChestPainType)
ggplot(data = heart_tbl, aes(x=RestingBP,y=MaxHR,group=HeartDisease,col=HeartDisease)) + geom_point()+ facet_wrap(~HeartDisease)
```
